Minimax-optimal regression over Sobolev Spaces with Neighborhood Graphs

### 2021-02-11 manuscripts/laplacian_smoothing/AISTATS/reviews/rebuttal.tex

\documentclass{article}

\usepackage{aistats2021_author_response}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % define colors in text
\usepackage{xspace}         % fix spacing around commands

\begin{document}

First of all, we would like to thank all the reviewers for their comments and helpful suggestions. We will begin with a general response that is relevant to multiple reviewers, and then give some specific responses to each reviewer.

\underline{General response} (\textit{``...limited experiment.''}) We agree that the theory would be better complemented by stronger empirical evidence. For a final version, we will extend our experiments to include different regression functions, higher dimensions, the power of our proposed test, and comparisons with other methods for nonparametric regression, especially with the thin-plate spline as suggested by Reviewer 4. We will also improve the presentation of our experimental results, as suggested by Reviewer 6.

\underline{Reviewer 1}
We are glad the reviewer enjoyed the paper.

\underline{Reviewer 2} (\textit{``...how large is d in practice?''}) In practice $d$ is often much larger than 4, while the intrinsic dimension $m$ is usually unknown. However, it is hypothesized (e.g in [Belkin 2006]) that $m$ is often small in real-world problems.

\underline{Reviewer 4}
(\textit{``...intuition behind the difference} [between Laplacian smoothing and thin-plate splines]...'') Laplacian smoothing penalizes the average size of the differences of the function evaluated at the design points, and thus insists on the estimate being on average ``smooth’’ across the design points. Thin-plate splines penalize the average size of the derivative across the domain, and thus are perfectly willing to accept estimates which are very ``wiggly'' near the design points, as long as the wiggliness is sufficiently localized. The key is that for sufficiently wiggly functions, derivatives and differences stop being coupled in high enough dimensions.

\underline{Reviewer 6} (\textit{``...this paper needs to make a more precise point as to its contributions to} [nonparametric regression] \textit{in particular...}) We study the method of Laplacian smoothing, which despite its long history has resisted analysis in the classical statistical setup. We show that although Laplacian smoothing is similar in spirit to first-order smoothing splines when $d = 1$, somewhat surprisingly it avoids the pathological failure of first-order thin-plate splines when $d \geq 2$. Furthermore, we establish that Laplacian smoothing adapts to the manifold hypothesis. Finally, we also show that Laplacian smoothing allows one to make only weak assumptions on “knowing the domain $\mathcal{X}$ and density $p$” while still obtaining minimax rates for Sobolev classes. We will make these points explicitly in our final version.

(\textit{“...a more precise comparison with other methods...”}) While it is possible to derive other minimax optimal regressors in our setup---e.g. by tuning kernel smoothing---our goal is somewhat different: to consider Laplacian smoothing, a widely used method with a strong empirical track record, and show that it specifically has optimal statistical properties. However, we agree with the reviewer that a more detailed comparison could be useful and will include this in the final version. Our analysis differs substantially from standard analyses of classical non-parametric methods (kernel smoothing, RKHS regression and truncated series estimators), and we note some important distinctions. \textit{Kernel smoothing}: the bias and variance terms we derive for Laplacian smoothing are quite different from those typical to kernel smoothing (even if the worst-case rates match), and indicate conditions under which Laplacian smoothing will outperform kernel smoothing. \textit{RKHS regression}: analyses of variational methods similar in spirit to Laplacian smoothing---i.e. thin-plate splines, but also more generally kernel ridge regression---typically assume that $f_0$ lies close to an RKHS; when $d \geq 2$ the first-order Sobolev space $H^{1}(\mathcal{X})$ is not an RKHS, and most standard analyses do not apply without careful modifications. \textit{Truncated series estimators}: the analyses of least squares over a subspace of $L^2(\mathcal{X})$ typically rely on an \textit{a priori} bound on the approximation error---which is hard to obtain in the manifold case without “knowing” the manifold in a strong sense.  In contrast, for Laplacian smoothing we derive a data-dependent bound on the approximation error, and upper bound this by a population-level quantity with high probability.

(\textit{“...practical implementation of the threshold.”}) In the final paper, we will make clear that the level-$\alpha$ permutation threshold is computed as follows: (1) randomly draw $M$ permutations $\pi_1,...,\pi_M$ of the indices $\{1,...,n\}$; (2) for each $m = 1,...,M$, compute the test statistic $T_m$ with respect to the permuted responses $(Y_{\pi_m(1)},X_1),...,(Y_{\pi_m(n)},X_n)$; (3) set the threshold to be the $(1 - \alpha)$-quantile of the empirical distribution of $\{T_1,...,T_M\}$.

(\textit{“...adaptivity to the intrinsic dimension m...”}) In our Theorems 3 and 4 the choice of radius $r$ depends on $m$. Typically, $m$ is unknown: although it can be estimated, practically we would recommend to choose $r$ by cross-validation. We believe, arguing along the lines of [Gyorfi 2006], that such a choice will achieve the same rates adaptively.

(\textit{“... intuition as to how your assumptions are used in your proofs?”}) In our final paper, we will include the following comments on assumptions. The Lipschitz assumptions, on the density and kernel, allow us to relate nonlocal approximations of the Sobolev semi-norm to the Sobolev semi-norm. The assumption that $K(c) > 0$ for some fixed $c > 0$ is fundamental---otherwise the radius $r$ does not truly reflect the scale of connectivity; on the other hand the requirement that the kernel be compactly supported on $[0,1]$ is for technical convenience. The lower bound on the density is used for comparing sample- and population-level eigenvalues; we make use of upper bounds on the optimal-transport distance in infinity norm between $P_n$ and $P$,  and such distance can be very large when the tails of $p(x)$ are close to $0$. Finally, the lower and upper bounds on the density are required for the lower and upper bounds in (16)---Weyl’s Law---to hold.

\end{document}

### 2021-02-11 manuscripts/laplacian_eigenmaps/laplacian_eigenmaps_regression.tex
Another round of re-structuring.

\begin{document}
\title{Minimax-optimal nonparametric regression with graph Laplacians}
\author{Alden Green}
\date{\today}
\maketitle

\section{Introduction}
\label{sec:introduction}

In the random design nonparametric regression problem, we observe data $(X_1,Y_1),\ldots,(X_n,Y_n)$, where $X_1,\ldots,X_n$ are independent samples from a distribution $P$ supported on a domain $\Xset \subset \Rd$, and 
\begin{equation}
\label{eqn:random_design_regression}
Y_i = f_0(X_i) + \varepsilon_i
\end{equation}
with $\varepsilon_i \sim N(0,1)$ independent Gaussian noise. Our goal is to perform statistical inference on the unknown regression function $f_0: \Xset \to \Reals$, by which we mean either (a) \emph{estimating} $f_0$ by $\wh{f}$, an estimator constructed from the data $(X_1,Y_1),\ldots,(X_n,Y_n)$ or (b) simply \emph{testing} whether $f_0 = 0$, i.e whether there is any signal present.

In graph-based nonparametric regression, we perform the above inferential tasks by first building a neighborhood graph $G_{n,r}$ which captures the geometry of $P$ and $\mc{X}$ in an appropriate sense, and then constructing an estimate $\wh{f}$ which is smooth with respect to the graph $G_{n,r}$. The neighborhood graph $G_{n,r} = ([n],{\bf W})$ is a weighted, undirected graph on vertices $[n] = \{1,...,n\}$, which we associate with the samples $\{X_1,\ldots,X_n\}$. The $n \times n$ weight matrix ${\bf W} = ({\bf W}_{ij})_{ij}$ encodes proximity between pairs of samples; for a kernel function $K: [0,\infty) \to \Reals$ and connectivity radius $r > 0$, the entries $\mathbf{W}_{ij}$ are given by
\begin{equation*}
\label{eqn:neighborhood_graph}
{\bf W}_{ij} = K\Biggl(\frac{\norm{X_i - X_j}_{\Rd}}{r}\Biggr).
\end{equation*}
Then the degree matrix ${\bf D}$ is the $n \times n$ diagonal matrix with entries ${\bf D}_{ii} = \sum_{j = 1}^{n}{\bf W}_{ij}$, and our star operator, the neighborhood graph Laplacian, can be written as
\begin{equation}
\label{eqn:graph_Laplacian}
\Lap_{n,r} = \bf{D} - \bf{W}
\end{equation}
The two graph-based regression methods we will consider are \emph{Laplacian eigenmaps} and \emph{Laplacian smoothing}. Each method uses $\Lap_{n,r}$ in a different manner to construct an estimator of $f_0$. Letting $v_1(G_{n,r}),\ldots,v_n(G_{n,r})$ denote eigenvectors of $\Lap_{n,r}$, the Laplacian eigenmaps estimator $\wh{f}_{\LE}$ \citep{belkin2003} is given by 
\begin{equation}
\label{eqn:laplacian_eigenmaps}
\wh{f}_{\textrm{LE}} = \sum_{k = 1}^{\kappa} \biggl\{\Bigl(\frac{1}{n}\sum_{i = 1}^{n} Y_i v_{k,i}(G_{n,r})\Bigr) v_{k}(G_{n,r})\biggr\}
\end{equation}
where $\kappa \in \{1,...,n\}$ is a tuning parameter. The Laplacian smoothing estimator $\wt{f}_{\LS}$ \citep{smola2003} is a penalized least squares estimator, given by
\begin{equation}
\label{eqn:laplacian_smoothing}
\wt{f}_{\LS}= \min_{f \in \Reals^n} \biggl\{\sum_{i = 1}^{n}(Y_i - f_i)^2 + \rho \cdot f^T \Lap_{n,r}^s f \biggr\}
\end{equation}
where $\rho > 0$ acts a tuning parameter on the penalty $f^T \Lap_{{n,r}}^s f$. Assuming~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are reasonable estimators of $f_0$, the squared empirical norms
\begin{equation}
T_{\LE} = \frac{1}{n}\Bigl\|\wh{f}_{\LE}\Bigr\|_2^2 \label{eqn:laplacian_eigenmaps_test}
\end{equation}
and
\begin{equation}
T_{\LS} = \frac{1}{n}\Bigl\|\wt{f}_{\LS}\Bigr\|_2^2 \label{eqn:laplacian_smoothing_test}
\end{equation}
are in turn reasonable statistics to test whether or not $f_0 = 0$.

Graph-based methods have a deserved place in the nonparametric regression toolbox. Of course, there exist a considerable range of methods for both the nonparametric testing and, especially, the nonparametric estimation problems. In fact, as we will see in Section~\ref{sec:problem_setup_and_background} some of these more classical methods are actually quite analogous to~\eqref{eqn:laplacian_eigenmaps}-\eqref{eqn:laplacian_smoothing}. Nevertheless, graph-based methods have many strengths: they are easy to implement; fast and stable to compute; generalizable to non-standard data---for instance, text or images, or really any data modality on which it is possible to define a kernel--- and non-standard learning paradigms---such as the semi-supervised setting. Last but not least, they have enjoyed widespread practical success.

For these reasons, a substantial body of work has arisen assessing the statistical properties of graph-based methods. Taken as a whole, this work views the graph Laplacian matrix as an estimator of a weighted continuum Laplace operator $\Delta_P$ (see equations~\eqref{eqn:laplace_operator} and~\eqref{eqn:laplace_operator_weighted} in Section~\ref{sec:problem_setup_and_background}), and establishes that when $n \to \infty$ and $r(n) \to 0$, it is in fact a \emph{consistent} estimator of $\Delta_P$. This notion of consistency is meant in a few different respects. For instance,  \citet{belkin03,belkin05} show that when $P$ is the uniform distribution over $\Xset$, $\Lap_{n,r}f \to \Delta f$ for sufficiently smooth functions $f$, \citet{lafon04,hein05} extend the statement to hold when $P$ is non-uniform, \citet{singer06} sharpens the convergence rates, and \cite{gine06} prove a central limit theorem. In another line of work, \citet{belkin07} prove that when $P$ is the uniform distribution over $\mathcal{X}$, the eigenvalues and eigenvectors of $\Lap_{n,r}$ converge to those of $\Delta_P$, \citet{garciatrillos18} extend this result to hold for when $P$, and \cite{calder2019} sharpen the rates at which $r(n)$ may converge to $0$.

These developments notwithstanding, the following fundamental theoretical question remains unanswered: are graph-based methods for nonparametric regression not merely consistent but in fact \emph{optimal}, in the minimax sense? In this paper, we confirm that they are. We establish that when the regression function $f_0$ is suitably smooth---by which we mean that it belongs to a first-order Sobolev ball; roughly speaking, that it possesses weak derivative $Df$ bounded in $L^2(\Xset)$ norm---both~\eqref{eqn:laplacian_eigenmaps} and~\eqref{eqn:laplacian_smoothing} are with high probability asymptotically rate-optimal estimators, and that both~\eqref{eqn:laplacian_eigenmaps_test} and~\eqref{eqn:laplacian_smoothing_test} are asymptotically rate-optimal test statistics. We also show that these estimators and test statistics can leverage additional assumed structure, either on the regression function $f_0$ or the distribution $P$. Our fundamental message is a reassuring one: the practitioner who wishes to use neighborhood graph procedures, for any of the various reasons outlined above, may do so without fear that they are sacrificing statistical optimality in the process.

\textcolor{red}{(TODO 1)}: Outline organization of paper here.

\paragraph{Notation.}
\begin{itemize}
	\item We write $\mathbb{Z}^d = \mathbb{Z} \otimes \cdots \otimes \mathbb{Z}$---with the tensor product being taken a total of $d$ times--- and likewise $\mathbb{N}^d = \mathbb{N} \otimes \cdots \otimes \mathbb{N}$. The sequence space $\ell^2(\mathbb{Z}^d)$ then contains those vectors $\theta: \mathbb{Z}^d \to \Reals$ for which $\sum_{k \in \mathbb{Z}^d} \theta_k^2 < \infty$. 
	\item For $1 \leq p < \infty$, we denote by $\Leb^p(\Xset)$ the set of functions $f:\Xset \to \Reals$ with bounded $p$th moment $\int_{\Xset} \bigl(f(x)\bigr)^p \,dx < \infty$. When $p = \infty$, we let $\Leb^{\infty}(\Xset)$ contain those functions $f$ for which $\sup_{x \in \Xset} |f(x)| < \infty$.
	\item For functions $f,g: \Xset \to \Reals$, we write $\dotp{f}{g}_n := \frac{1}{n}\sum_{i = 1}^{n} f(X_i) g(X_i)$, which we refer to as the \emph{empirical inner product}, and $\norm{f}_n^2 = \dotp{f}{f}_n$, which we refer to as the (squared) \emph{empirical norm}. For vectors $v \in \Reals^n$, we write $\norm{v}_p^p := \sum_{i = 1}^{n} \abs{v_i}^p$. When there is no chance of being confused, we will sometimes associate a function $f$ with its evaluations $(f(X_1),\ldots,f(X_n)) \in \Reals^n$.
	\item In general, for a graph $G$ we use $\Lap_G$ to denote the Laplacian of $G$. The matrix $\Lap_G$ is a square, symmetric, positive semi-definite matrix, with $n$ eigenvalues $\lambda_k(G)$ and eigenvector $v_k(G)$ pairs, defined according to
	\begin{equation*}
	\Lap_G v_k(G) = \lambda_k(G) v_k(G), ~~\frac{1}{n} \norm{v_k(G)}_2^2 = 1 
	\end{equation*}
	and ordered so that $0 \leq \lambda_1(G) \leq \ldots \leq \lambda_n(G)$. As we've already seen, to reduce notational clutter we write $\Lap_{n,r}$ instead of $\Lap_{G_{n,r}}$. 
	\item For sequences $a_n, b_n$, we write $a_n \lesssim b_n$ when there exists a constant $C > 0$ such that $a_n \leq C b_n$ for all $n$, and $a_n \asymp b_n$ when $a_n \lesssim b_n$ and $b_n \lesssim a_n$. 
	\item When convenient we write $\mathbf{X} = (X_1,\ldots,X_n)$. 
\end{itemize}

\section{Problem Setup and Background}
\label{sec:problem_setup_and_background}

Before we get to our main results, we will review some of the relevant literature on minimax optimal estimation and testing of Sobolev functions. We will also recall some (of the many) estimators and tests which achieve these rates: in particular the \emph{Fourier series projection} and \emph{smoothing spline} methods. As we will see, one can interpret Laplacian eigenmaps and Laplacian smoothing as graph-based, i.e. discrete, counterparts to these ``continuum'' methods.

\subsection{Sobolev spaces}
\label{subsec:sobolev_spaces}

We start by briefly introducing Sobolev spaces, emphasizing only the material relevant to our statistical context. Roughly speaking, Sobolev spaces contain functions $f \in \Leb^p(\Xset)$ with derivatives which themselves belong to $\Leb^p(\Xset)$. More formally, for given integers $s$ and $p > 0$, the Sobolev space $W^{s,p}(\Xset)$ consists of all functions $f \in \Leb^p(\Xset)$ such that for each multiindex $\alpha = (\alpha_1,\ldots,\alpha_d) \in \mathbb{N}^d$ satisfying $\abs{\alpha} := \sum_{i = 1}^{d} \alpha_i \leq s$, the weak derivative $D^{\alpha}f$ exists and belongs to $\mathcal{L}^p(\Xset)$. For such functions, the $(s,p)$-Sobolev seminorm $\seminorm{f}_{W^{s,p}(\Xset)}$ and norm $\norm{f}_{W^{s,p(\Xset)}}$ are given by 
\begin{equation*}
\seminorm{f}_{W^{s,p}(\Xset)}^p = \sum_{\abs{\alpha} = s}\int_{\mathcal{X}} \Bigl|\bigl(D^{\alpha}f\bigr)(x)\Bigr|^p \,dx, ~~ \norm{f}_{W^{s,p}(\Xset)}^p = \sum_{k = 0}^{s} \seminorm{f}_{W^{k,p}(\Xset)}^p
\end{equation*}
and for a given $M > 0$, the $(s,p)$-Sobolev ball is $W^{s,p}(\Xset, M) = \set{f: \norm{f}_{W^{s,p}(\Xset)} \leq M}$. In the special case when $p = 2$, the Sobolev space $W$ is a Hilbert space, and we adopt the usual convention of writing $H^s(\Xset) = W^{s,p}(\Xset)$ and $H^s(\Xset,M) = W^{s,2}(\Xset,M)$; we will confine our attention to this case hereafter.

Periodic functions $f \in H^s(\Xset)$ admit a representation in terms of Fourier coefficients. Define the \emph{Sobolev ellipsoids} $\Theta_s(M) \subseteq \ell^2(\mathbb{Z}^d)$ to be
\begin{equation*}
\Theta_s(M) = \Bigl\{\theta \in \ell_2(\mathbb{Z}^d): \sum_{k \in \mathbb{Z}^d} \theta_k^2 \abs{k}^{2s} \leq M^2\Bigr\}.
\end{equation*}
To see the connection between $H^s(M)$ and $\Theta_s(M)$, for the moment let us fix the domain $\Xset = [0,1]^d$, and write $\set{\phi_k}_{k \in \mathbb{Z}^d}$ for the trigonometric basis of $\Leb^2(\Xset)$. Then the periodic Sobolev balls
\begin{equation*}
H_{\textrm{per}}^s(\Xset,M) = \Bigl\{\sum_{k \in \mathbb{Z}^d} \theta_k \phi_k: \theta_k \in \Theta_s(M) \Bigr\}
\end{equation*}
satisfy $H_{\textrm{per}}^s(\Xset,M) \subseteq H^s(\Xset,M_1)$ for a sufficiently large value of $M_1 > 0$. On the other hand, any $1$-periodic function $f \in H^s(\Xset,M)$ also belongs to $H_{\textrm{per}}^s(\Xset,M_0)$ for a sufficiently large value of $M_0 > 0$.

We will also be interested in the \emph{zero-trace} Sobolev spaces, which impose more severe boundary conditions than periodicity. Letting $\partial \Xset$ denote the boundary of $\Xset$, informally one can think of the order-$s$ zero-trace Sobolev space as the set of functions $f \in H^s(\Xset)$ for which ``$f(x) = 0$ for all $x \in \partial\Xset$''. However, since functions $f \in H^s(\Xset)$ are defined only up to sets of measure zero, such a definition is not really meaningful. Instead, letting $C_c^{\infty}(\Xset)$ consist of those functions in $C^{\infty}(\Xset)$ which are compactly supported in $\Xset$, we formally define the order-$s$ zero-trace Sobolev space $H_0^{s}(\Xset)$ as the completion of $C_c^{\infty}(\Xset)$ in the $H^{s}(\Xset)$ norm. When the domain $\Xset = [0,1]^d$, it holds that $H_0^{s}(\Xset) \subseteq H_{\textrm{per}}^{s}(\Xset)$. \textcolor{red}{Justify this statement with a reference.}

\textcolor{red}{(TODO 1)}: Introduce the Sobolev embedding theorem, for at least three reasons: (1) to facilitate later discussion of why smoothing splines are ill-posed when $2s < d$; (2) to make explicit why the statistical problem isn't ill-posed when $2s < d$ in the random design setting; (3) to make clear why the testing problem is hopeless when $4s < d$. Conclude with one special fact worth mentioning: regardless of $s$ and $d$, when $\Xset$ is sufficiently smooth for any Sobolev spaces $H^s(\Xset)$ the set of infinitely differentiable functions $C^{\infty}(\Xset)$ is dense in $H^s(\Xset)$. Said differently, for each integer $s$, the Sobolev space $H^s(\Xset)$ is simply the completion of $C^{\infty}(\Xset)$ in the norm $\norm{\cdot}_{H^s(\Xset)}$.

\subsection{Nonparametric regression: Fourier series and smoothing splines}
\label{subsec:continuum_methods}
Orthogonal projection onto the Fourier basis and smoothing splines are two approaches particularly well-suited for nonparametric regression over Sobolev spaces. In the former approach, we estimate the Fourier coefficients $\wh{\theta}_k := \dotp{Y}{\phi_k}_n$, with the resulting function estimate
\begin{equation}
\label{eqn:orthogonal_series}
\wh{f}_{\OS} = \sum_{k: \abs{k} \leq \kappa} \wh{\theta}_k \phi_k
\end{equation}
for some $\kappa \geq 0$. In contrast, the smoothing spline estimator $\wt{f}_{\SM}$ is a variational estimator, defined as the solution to the following optimization problem,
\begin{equation}
\label{eqn:smoothing_spline}
\wt{f}_{\SM} = \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \seminorm{f}_{H^{s}(\Xset)}^2\biggr\}
\end{equation}
with the minimum taken over all $f \in H^s(\Xset)$.

The orthogonal series estimator $\wh{f}_{\OS}$ is closely connected with the Laplacian eigenmaps estimator $\wh{f}_{\LE}$. Just as $v_k(G_{n,r})$ are eigenvectors of the graph Laplacian $\Lap_{n,r}$, so too are $\phi_k$ eigenfunctions of a particular differential operator $\Delta: C^2(\Xset) \to \Leb^2(\Xset)$, known as the \emph{Laplace} operator and formally defined as
\begin{equation}
\label{eqn:laplace_operator}
\Delta g := -\mathrm{div}(\nabla g),
\end{equation}
where $\mathrm{div}(\cdot)$ stands for the divergence, and $\nabla$ for the gradient. As mentioned previously, when the radius $r \to 0$ as $n \to \infty$ at an appropriate rate, the graph Laplacian $\Lap_{n,r}$ in fact converges to a weighted Laplace operator $\Delta_P$, given by
\begin{equation}
\label{eqn:laplace_operator_weighted}
\Delta_{P} g := -\frac{1}{p}\mathrm{div}(p^2 \nabla g);
\end{equation}
likewise the eigenvectors $v_k(G_{n,r})$ converge to eigenfunctions of $\Delta_{P}$. The upshot is that when $P$ is the uniform measure on $\Xset = [0,1]^d$---and thus the eigenfunctions of $\Delta_P = \Delta$ are precisely the trigonometric basis functions $\phi_k$--- and $\kappa \geq 0$ is fixed, the estimator $\wh{f}_{\OS}$ is in fact the limit of $\wh{f}_{\LE}$, in the sense that
\begin{equation}
\label{eqn:laplacian_eigenmaps_consistency}
\abs{\wh{f}_{\LE}(X_i) - \wh{f}_{\OS}(X_i)} \overset{P}{\to} 0~~\textrm{as $n \to \infty$ and $r(n) \to 0$,}
\end{equation} 
for each $i \in \mathbb{N}$.

By viewing $\Delta_P$ as the continuum limit of~$\Lap_{n,r}$, one can also tie the penalty in the smoothing spline estimator $\wh{f}_{\SM}$ to that of the Laplacian smoothing estimator $\wh{f}_{\LS}$. Using the definition~\eqref{eqn:laplace_operator}, repeated integrating by parts, and then invoking the aforementioned convergence properties of $\Lap_{n,r}$, it can be shown that for any $f \in C^{2s}(\Xset)$ (see \citet{zhou11} for an argument in this spirit),
\begin{equation}
\label{eqn:pointwise_seminorm_convergence}
\abs{f}_{H^s(\Xset)} = \int_{\Xset} \bigl(\Delta_P^sf\bigr)(x) \cdot f(x) \,dx = \lim_{n \to \infty} \frac{1}{n^{s + 1}r^{s(d + 2)}}f^T \Lap_{n,r}^s f.
\end{equation}
This does not imply a formal statement about the convergence of $\wh{f}_{\LS}$ to $\wh{f}_{\SM}$ along the lines of~\eqref{eqn:laplacian_eigenmaps_consistency}. Among other issues, the convergence in~\eqref{eqn:pointwise_seminorm_convergence} is over $f \in C^{2s}(\Xset)$ while the domain of minimization in~\eqref{eqn:smoothing_spline} is $H^{s}(\Xset)$. Nevertheless, it establishes a general connection between $\wh{f}_{\LS}$ and $\wh{f}_{\SM}$.

\subsection{Minimax rates over Sobolev classes}
\label{subsec:minimax_rates}
One reason all of this is intriguing is that, as mentioned previously, the continuum estimators $\wh{f}_{\OS}$ and $\wh{f}_{\SM}$---and test statistics based on functionals of these estimators--- have strong statistical properties. In particular, these estimators and tests are known to have error rates within a constant factor of minimax optimal. Throughout this section, we treat $M > 0$ as a fixed number (i.e. not varying with $n$.)

\paragraph{Estimation rates.}
The minimax estimation rate over Sobolev balls is 
\begin{equation}
\label{eqn:sobolev_space_estimation_minimax_rate}
\inf_{\wh{f}} \sup_{f_0 \in \mc{H}^s(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d \geq 0$,}
\end{equation}
where the function class $\mc{H}^s(\Xset,M)$ is any one of $H^s(\Xset,M)$, $H_{\mathrm{per}}^s(\Xset,M)$ or $H_0^s(\Xset,M)$.

Fourier series and smoothing spline estimators are two (of the many) estimators which certify the upper bound in~\eqref{eqn:sobolev_space_estimation_minimax_rate}. More specifically, if we assume that
\begin{enumerate}[label=(P\arabic*)]
	\item
	\label{asmp:uniform_density}
	the distribution $P$ is the uniform probability measure on the unit cube $\Xset = [0,1]^d$,
\end{enumerate}
then $\wh{f}_{\OS}$ is minimax optimal over the Sobolev ellipsoids $H_{\mathrm{per}}^s(\Xset, M)$, meaning
\begin{equation}
\label{eqn:fourier_series_estimation_rate}
\sup_{f_0 \in H_{\mathrm{per}}^s(\Xset,M)}\Ebb\Bigl[\norm{\wh{f}_{\OS} - f_0}_{L^2(\Xset)}^2\Bigr]  \lesssim n^{-2s/(2s + d)}~~\textrm{for all $s$ and $d \geq 0$.}
\end{equation}
In contrast, if we merely assume that
\begin{enumerate}[label=(P\arabic*)]
	\setcounter{enumi}{1}
	\item 
	\label{asmp:bounded_density}
	The density $p$ is bounded away from $0$ and $\infty$, that is
	\begin{equation*}
	0 < p_{\min} \leq p(x) \leq p_{\max} < \infty
	\end{equation*}
	for all $x \in \Xset$,
\end{enumerate}
then $\wt{f}_{\SM}$ is minimax optimal over the Sobolev balls $H^s(\Xset,M)$, meaning
\begin{equation}
\label{eqn:smoothing_spline_estimation_rate}
\sup_{f_0 \in H^s(\Xset,M)}\Ebb\Bigl[\norm{\wh{f}_{\SM} - f_0}_{L^2(\Xset)}^2\Bigr]  \lesssim n^{-2s/(2s + d)}~~\textrm{for all $2s > d$.}
\end{equation}
These results reveal some relative strengths and weaknesses of $\wh{f}_{\OS}$ and $\wh{f}_{\SM}$. On the one hand, the result for smoothing splines clearly invokes weaker regularity conditions than the result for Fourier series projection. The assumption~\ref{asmp:bounded_density} is less stringent than~\ref{asmp:uniform_density}, in that it does not specify that either the density $p$ or domain $\Xset$ be known. Moreover,~\eqref{eqn:fourier_series_estimation_rate} requires that the regression function $f_0$ be periodic, whereas~\eqref{eqn:smoothing_spline_estimation_rate} does not. On the other hand, the restriction $2s > d$ in~\eqref{eqn:smoothing_spline_estimation_rate} is a severe limitation of smoothing splines relative to Fourier series estimators. It is worth emphasizing that this is a fundamental defect of smoothing splines, as opposed to any flaw in the theory; when $2s \leq d$ the optimization problem~\eqref{eqn:smoothing_spline} is not even well-defined! Interestingly, many of these points will reappear when we turn to graph-based estimators. To be explicit: we will see that the Laplacian smoother $\wh{f}_{\LS}$ fails when $d$ is too large relative to $s$; that our analysis of $\wh{f}_{\LE}$ requires that $f_0$ be periodic when $s > 1$; and perhaps most notably, that neither the estimators nor our analysis requires $p$ or $\Xset$ to be known.

\textcolor{red}{(TODO 2):} succinctly explain why smoothing splines are ill-posed when $2s \leq d$, by referring to the Sobolev Embedding Theorem.

\paragraph{Goodness-of-fit testing rates.}

In the goodness-of-fit testing problem, we ask for a test function (formally, a Borel measurable map $\phi: \mc{X}^n \times \Reals^n \to \{0,1\}$) which can distinguish between the null and alternative hypotheses
\begin{equation}
\mathbf{H}_0: f_0 = 0, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \neq 0.
\end{equation}
We measure the performance of a test $\phi$ by its Type I error $\Ebb_0[\phi]$, and Type II error $\mathbb{E}_{f_0}[1 - \phi]$. It is well known that no test $\phi$ with controlled Type I error---say, $\Ebb_0[\phi] \leq 1/b$ for some constant $b > 1$---
can have controlled Type II error---$\Ebb_{f_0}[\phi] \leq 1/b$--- for every $f_0 \neq 0$. As a result, it is standard to impose two kinds of restrictions on the alternative hypothesis class $\mathbf{H}_a$. The first is a smoothness condition on $f_0$: in our case, we will assume that $f_0$ belongs to a Sobolev ball $H^s(\Xset,M)$. The latter is a separation condition enforcing a minimum distance between $f_0$ and $0$: in our case, we will insist that $\norm{f_0  - 0}_{\Leb^2(\Xset)} > \epsilon$ for some $\epsilon > 0$.

Once smoothness and separation conditions are established, the following becomes a natural question to ask: what is the smallest value of $\epsilon$ for which there exists a test $\phi$ with both small Type I error, and small Type II error over all functions $f_0$ which are smooth and separated from $0$ by at least $\epsilon$? This smallest separation $\epsilon$ is known as the critical radius, and in the case of the Sobolev classes it is given by
\begin{equation}
\label{eqn:sobolev_space_testing_critical_radius}
\epsilon_n^{\star}\Bigl(\mc{H}^s(\Xset,M)\Bigr):= \inf\biggl\{\epsilon: \inf_{\phi} \Bigl\{\Ebb_0[\phi] +  \sup_{\substack{f_0 \in \mc{H}^s(\Xset,M), \\ \norm{f_0}_{\Leb^2(\Xset)} \geq \epsilon}} \Ebb_{f_0}[1 - \phi]\Bigr\} \leq \frac{2}{b} \biggr\} \asymp n^{-2s/(4s + d)}~~\textrm{for all $4s > d$,}
\end{equation}
where $\mc{H}^s(\Xset,M)$ is any one of $H^s(\Xset,M)$, $H_{\mathrm{per}}^s(\Xset,M)$ or $H_0^s(\Xset,M)$.  Just as $\wh{f}_{\OS}$ and $\wh{f}_{\SM}$ achieve minimax optimal estimation rates, tests using the plug-in statistics
\begin{equation*}
T_{\OS} = \bigl\|\wh{f}_{\OS}\bigr\|_{\Leb^2(\Xset)}^2,~~\textrm{and}~~T_{\SM} = \bigl\|\wt{f}_{\SM}\bigr\|_{n}^2
\end{equation*}
are minimax rate-optimal for the goodness-of-fit testing problem over Sobolev balls; see \citet{ingster2009} for analysis of the former statistic, and \citet{liu2019} for analysis of the latter (when $2s > d$).

When $4s \leq d$, the Sobolev classes $H^s(\Xset,M)$ are too irregular for any test statistic to accurately discern whether $\norm{f_0}_{\Leb^2(\Xset)} = 0$; in particular, some functions in these classes will not have a bounded $4$th moment. However, we will show in Section~\ref{sec:minimax_optimal_graph_Laplacian_methods} that upon explicitly assuming a bounded 4th moment, one obtains the dimension-free minimax rate
\begin{equation*}
\epsilon_n^{\star}\Bigl(\Leb^4(\mc{X},M)\Bigr) \asymp n^{-1/4},
\end{equation*}
revealing a distinction between the estimation and goodness-of-fit testing problems. This extends the results of \cite{guerre02}, who show the analogous result over Holder balls.

\paragraph{Manifold hypothesis.}

Suppose that $\Xset$ satisfies the manifold hypothesis, meaning informally that it has intrinsic dimension $m < d$ (see~\ref{asmp:domain_manifold} in Section~\ref{sec:graph_sobolev_classes} for a formal definition). In this context, both~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$. What's more, suitable adaptations of the estimators $\wh{f}_{\OS}$ and $\wt{f}_{\SM}$, and the test statistics $T_{\OS}$ and $T_{\SM}$, achieve these faster rates. We will not further review the details here, except to emphasize one point: unlike in the full-dimensional context, in order for either the smoothing spline or Fourier series approaches to optimally leverage the manifold hypothesis, the domain $\Xset$ must be known. In contrast, we will see that our graph-based procedures will be optimal without this knowledge; in this sense, they automatically ``feel'' the geometry of $\Xset$. We will return to this point in Section~\ref{sec:discussion}.

\subsection{Our contributions}
\label{subsec:our_contributions}

>>>>

\begin{document}
\title{Minimax-optimal regression over Sobolev Spaces with Neighborhood Graphs}
\author{Alden Green}
\date{\today}
\maketitle

\bibliographystyle{plainnat}
\bibliography{../../graph_testing_bibliography}
